Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset

### 2024-10-11 code/manuscript.Rmd
Added references and removed continue sampling from result sections

---
title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA] ^[Correspondence to aoyuzou@berkeley.edu], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
abstract: Conventional measurement and verification (M&V) methods for whole-building energy savings estimation are both time-consuming and unreliable, especially when non-routine events occur during the M&V process. Those events are unrelated to the proposed intervention strategy but have substaintial impacts on the building energy consumption. In this study, we argue that for switchable interventions (e.g. most of the control retrofits) can benefit from random sampling where the analyst randomly decide which strategy (i.e. baseline or intervention) to implement each day. We tested the novel randomized M&V method on a large public dataset which covers multiple climate zones and types of commercial buildings. We applied a virtual chilled water supply temperature reset based on outdoor weather as a control retrofit intervention. Our study shows that the new M&V method can estimate the savings accurately much quicker than the conventional method and most importantly, the estimation results are much more robust compared to the conventional method when non-routine events are present.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
# bibliography: references.bib
# biblio-style: apalike

@@

The conventional M&V process can be found in ASHRAE Guideline 14
<!--# TODO:Add reference -->, which are summarized in Figure
\@ref(fig:convmethod). The process normally starting with baseline
measurement for a whole year before retrofit. After deploying the
intervention, building performance measurement continues for another
year. At the end of the two-year period,

@@

### Non-routine events

A common non-routine event in an energy-saving M&V project is a change
in occupancy or a substantial change in occupant behavior. Those changes
have a significant impact on measured building energy consumption and
are typically not caused by the intervention strategy. If the M&V
analyst are not aware of the change and has no reasonable approach to
make adjustment on the energy measurements, the saving estimation result
is largely biased. For example, most commercial buildings are unoccupied
during the pandemic in 2020 and thus building managers have observed a
drastically decrease in monthly energy bills despite no energy-efficient
measures were implemented. In addition, common adjustments by an
building analyst only consider the variations of outdoor weather
conditions such as by fitting a regression model (e.g. Time-Of-Week
Temperature model).

@@

@@

### Measurement and verification

### Randomized experimental design

@@

As mentioned, the goal of a M&V project is to determine the effect
(normally savings) of an energy-efficient intervention. And in this
study, we further limit the study scope to switchable interventions,
which mostly encapsulates control retrofits. An example intervention of
such type can be a control retrofit developed by a software-as-a-service
company that enables the chilled water plant to reset its supply water
temperature based on outdoor weather condition. Therefore, we defined
the M&V scenario as follows:

*"A company wishes to sell their supply temperature reset control
software package to a building owner and guarantees its effect in
reducing building electricity usage. If the building owner decides to
purchase the service, the company agrees to charge the service fee based
on a fraction of the measured savings."*

As required by the M&V scenario, we assessed the performance of both the
conventional and the novel randomized M&V methods by estimating the
intervention energy savings for all valid buildings in the dataset. By
conducting such analysis, we hope to:

1.  Demonstrate the implementation of the proposed randomized M&V method
    using a public available dataset. We ensured the reproducibility of
    the M&V method by making the analysis code open source including
    randomized schedule generation, sequential statistical analysis,
    energy modeling and normalized saving calculation. Using the
    available open resources, building analysts should be able to
    seamlessly integrate and apply them in their own M&V projects.

2.  Compare the energy saving estimation accuracy between the
    conventional and the randomized method. In particular, this study
    extends the comparison to large samples of buildings of various
    types and across multiple climate zones.

3.  Verify the superior robustness of the randomized method over the
    conventional method. By using the realistic measurements from
    real-world buildings, which contains various sources of noises,
    could largely reflect the challenges that a building analyst would
    be facing in any real project. Particularly, as we will demonstrate
    in the following sections, non-routine events (i.e. 'noises' in the
    measurements) have less impact in the energy saving estimation when
    the analyst uses the randomized approach.

@@

As mentioned, this study leverages a large public dataset to demonstrate
the energy saving estimation results of a novel M&V method inspired by
other scientific research fields. We outlined the methodology of the
study in Figure \@ref(fig:flowchart) and outlined several key components
in this section.

@@

1.  Missing values \< 1000: given the hourly resolution of all
    measurements, this is equivalent to more than a month of missing
    days.

@@

4.  Target buildings should have known site location: buildings with
    anonymous location can not request typical meteorological weather.

5.  Warehouse and parking types are excluded: target buildings that are
    likely to have non-regular electricity and chilled water usage
    should be excluded.

As a summary, the resulting subset contains all buildings with 'tidy'
measurements. Figure \@ref(fig:tidysite) shows in total, the subset
contains `r nrow(all_names_tidy)` buildings in `r nrow(all_types_tidy)`
types from `r nrow(all_sites_tidy)` different climate zones.

@@

The 'tidy' subset might be presentative to the measurements collected
among the existing building stock since in reality, whole-building
electricity measurements collected over two years normally exhibit
larger changes. Those measured changes can be caused by sensor itself
such as lack of calibration, or inherent changes of the building, which
illustrated before, can bias the result of a M&V. Therefore, to compare
the robustness of the two M&V method more realistically, we included an
additional 'messy' subset which first exclude the 'tidy' subset and then
re-apply the filtering rule with one amendment:

3.  Absolute mean difference of the two-year electricity usage \< 25%:
    any increase or decrease of building electricity usage in the second
    year should be less than 25% of that in th first year.

Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset. As
this is less aggressive, it contains `r nrow(all_names_messy)` buildings
in `r nrow(all_types_messy)` types from `r nrow(all_sites_messy)`
different climate zones.

@@

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions. Under both strategies, we assume the chiller
is enabled when outdoor temperature is higher than 10°C. The baseline,
which assumed to be the existing measurements from the dataset, runs at
a constant water supply temperature. The intervention as the figure
shows reset from 7 °C to 12 °C.

@@

We map the chilled water supply temperature reset to the electrical
energy savings as: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

We assume in general, HVAC uses around 50% of whole-building electricity
and the chilled water plant further uses 50% electricity of the HVAC
system. This assumption largely simplifies the unique characteristics of
electricity usage in all types of commercial buildings, but for the
scope of this paper, we therefore assume 25% of whole-building
electricity is consumed by the chilled water plant ($E_{CHW}$).
<!--# TODO: Add reference -->In most cases, savings from an intervention
is unproportional to the building hourly electricity usage, which
renders M&V challenging. Therefore, we map the resulting electricity
savings as a percentage of the mean electricity consumption of the plant
determined by outdoor temperature ($OAT$), intervention supply water
temperature ($T_{interv}$), baseline supply water temperature
($T_{base}$) and hour of the day ($\delta_{occ}$, binary indicator
whether during peak hours from 9 AM to 4 PM).

@@

Parameters and their pre-defined values are summarized in the table
below. Those parameters were not rigorously calibrated for each
building, and were used uniformly across the dataset. Although this
calibration process should be included, it is considered tangential to
the study scope outlined in this paper.

| Parameter              | Description                                                      | Value |
|------------------|------------------------------------|------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

: Parameters for calculating the intervention savings

@@

To demonstrate the influence of non-routine events that are not properly
adjusted, we developed three scenarios to quantify the energy
consumption change associated with a hypothetical occupancy change. In
summary, we assume an increase in occupancy in 2016 (i.e. refered as the
'baseline year' if using conventional method) resulted in 20% increase
in whole-building measured electricity usage. The three scenarios
specifies the increase happened during January to April (S1); May to
August (S2); and September to December (S3). We applied such change only
to the buildings in the 'tidy' subset as originally they are more likely
to have no non-routine events.

@@

```{r convmethod, fig.cap="Flow chart showing the conventional M&V process for an energy-saving intervention", out.width="50%"}

knitr::include_graphics(paste0(fig_path, "convmethod.png"))
```

As described in Section \@ref(introduction) and Figure
\@ref(fig:convmethod), conventional method for M&V is time-consuming and
unable to separate non-routine event impact from measured savings. If
using conventional method to estimate intervention energy savings, the
result (after 24 months) is calculated by the difference between
measured intervention and projected baseline in the post-retrofit
period. In this study, we leveraged a piece-wise linear regression
considering time-of-week and outdoor temperature (TOWT) as independent
variables for projection and normalization on typical meteorological
year <!--# TODO: Add reference -->.

@@

```{r randmethod, fig.cap="Flow chart showing the novel randomized M&V process for an energy-saving intervention", out.width="50%"}

knitr::include_graphics(paste0(fig_path, "randmethod.png"))
```

Compared to the conventional method, the randomized M&V can provide an
estimation more rapid and reliable. Section \@ref(introduction)
describes the application of randomization techniques and Figure
\@ref(fig:randmethod) specifically outlined the process of conducting a
randomized switchback experiment for control retrofit projects. The
method sequentially evaluate the intervention effect and returns a
saving estimation with 95% confidence interval when all stopping
criteria are satisfied. <!--# TODO: Add paper DOI --> describes all
stopping criteria and recommened threshold to consider in more details
and we breifly summarized the thresholds used in this paper here:

@@

The timeline for the conventional M&V method is outlined in previous
sections. Although ASHRAE Guideline 14 offers minimum requirements for
each path, the required baseline period is either a full range of all
independent variables (typically outdoor weather conditions) under
normal facility operation or a 12-month worth of continuous
measurements. The same requirements also applies to intervention
installed in the post-retrofit period. Thus, the conventional M&V method
is likely to run over 24 months or even longer due to missing data. For
example if an analyst is asked to determine the energy savings of a
chilled water plant retrofit but missed most of the cooling season due
to delay in retrofit deployment, he/she needs to wait until the next
cooling season to measure the savings.

If the randomized M&V method is applicable, even if there is a delay in
the retrofit deployment, the building owner faces less risk and impact
since random sampling allows measuring the full range of all independent
variables simultaneously among all control strategies or modes. In
addition, sampling at an equal probability (i.e. 50%/50% between
baseline/intervention) effectively balances the level of independent
variables measured. In other words, unless one strategy is sampled with
a large amount of consecutive days (e.g. 7 days or more), it is rare
that the independent variable, such as outdoor weather condition, is
measured significantly different among all sampled control strategies.
Figure \@ref(fig:timeline) shows the average estimated timeline for each
building across all sites and climate conditions if using randomized M&V
for the applied chilled water supply temperature reset intervention. The
timeline figures for each individual building are attached in the
supplementary material. The figure shows for all climate zones, covering
a sufficient range outdoor weather condition is the most stringent
criterion. However, most buildings can satisfy the 80% range by 6 months
and all buildings can quantitatively determine the savings with
associated uncertainty by 9 months, which is even shorter than the
length of baseline measurement required by the conventional method.

@@

In this study, we define the overall target savings (i.e. the ground
truth) that an M&V analyst wishes to detect, either in normalized
savings, fractional savings or simply the measured difference between
the baseline and the intervention measurements, is the mean electricity
energy reduction over the two-year period from implementing the chilled
water supply temperature reset control. We estimate the M&V savings by
the two methods following the process given in Figure
\@ref(fig:convmethod) and Figure \@ref(fig:randmethod). In orther words,
we compare the conventional M&V savings at the end of the two-year
period with the randomized M&V savings after all criteria are satisfied.

Figure \@ref(fig:norm) shows the overall comparison of the savings
estimation from all buildings normalized on the typical meteorological
year weather conditions of each site. In subplot a), the narrower range
of the true savings indicates the dependence of the intervention effect
on the outdoor weather condition, which is intended as shown in Figure
\@ref(fig:chwst). Sites with mild climate all year round such as
locations in California shows higher savings potential above 10%, while
locations with more extreme climate such as Washington DC shows only 6%
savings annually. In addition, although the box plot indicates some
savings estimation are 'outliers', they should not be interpreted as
such since the distribution can be biased by the unbalanced sample sizes
across different site shown in Figure \@ref(fig:tidysite) and Figure
\@ref(fig:messysite). In subplot b), the results show that the
conventional M&V method tends to estimate savings with a greater
uncertainty and we will demonstrate in the following sections that
non-routine events can have a significant impact. For the randomized
method, we displayed two scenarios, the estimation at the end of
two-year period with 50%/50% sampling throughout and the estimation
provided after all stopping criteria satisfied (i.e. stops early at 24
weeks or 36 weeks). Since we found no significant difference between the
two distributions as shown in the figure, we demonstrated that an early
stop has no impact on the estimation accuracy but can significantly
reduces the cost associated with M&V.

@@

p1 <- df_FS_tidy %>% 
  bind_rows(df_FS_messy) %>% 
  filter(method == "true") %>% 
  mutate(method = "true savings") %>% 
  ggplot(aes(x = method, y = savings, fill = method)) +
  stat_boxplot(geom ='errorbar', width = 0.5) +
  geom_boxplot(outlier.size = 1) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     limits = c(0, 30), 
                     labels = number_format(suffix = "%")) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

>>>>

---
title: "Reliability Demonstration of A Novel Randomized Measurement and Verification Method for Switchable Control Retrofit Using Large-scale Public Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Durate ^1^
abstract: Conventional measurement and verification (M&V) methods for whole-building energy savings estimation are both time-consuming and unreliable, especially when non-routine events occur during the M&V process. Those events are unrelated to the proposed intervention strategy but have substaintial impacts on the building energy consumption. In this study, we argue that for switchable interventions (e.g. most of the control retrofits) can benefit from random sampling where the analyst randomly decide which strategy (i.e. baseline or intervention) to implement each day. We tested the novel randomized M&V method on a large public dataset which covers multiple climate zones and types of commercial buildings. We applied a virtual chilled water supply temperature reset based on outdoor weather as a control retrofit intervention. Our study shows that the new M&V method can estimate the savings accurately much quicker than the conventional method and most importantly, the estimation results are much more robust compared to the conventional method when non-routine events are present.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

@@

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008,
@efficiency_valuation_organisation_international_2007,
@ashrae_ashrae_2014]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level where the measurements are obtained from utility
bills or whole-building meters. The corresponding M&V process required
by ASHRAE Guideline 14, which we refer to as the convention method in
this paper. Typically, the process begins with baseline measurements
taken over a year before implementing any energy-efficiency retrofit,
followed by the same measurement procedure during the intervention
period. After collecting two years of data, an M&V analyst fits an
energy prediction model, using variables such as outdoor temperature and
time [@mathieu_quantifying_2011] to project baseline energy consumption
in the post-retrofit period. The difference between the counter-factual
baseline and the measured intervention represents the energy savings. A
key drawback of this method is its reliance on a two-year timeline to
quantify savings, during which baseline measurements can become outdated
due to changes in building performance caused by non-routine events
unrelated to the intervention. This limitation reduces the feasibility
of rapid M&V and complicates the quantification of estimation
uncertainty, thus impacting the accuracy and timeliness of savings
assessments.

@@

To address the limitations of conventional M&V methods and the
challenges posed by non-routine events, we propose a novel M&V method
that adopts the randomized crossover design, a concept borrowed from
medical and agricultural studies. Another improvement is that we
proposed a sequential evaluation framework and defined stopping criteria
to end the M&V if the target effect is detected. This is to avoid
unnecessary measurement collection over the full 2-year M&V cycle. The
full framework is detailed in another study with all stopping criteria
outlined <!--# TODO: Add references -->. In summary, this method
provides M&V analysts with a randomized schedule that alternates between
baseline and intervention implementation while ensuring balanced
sampling across days of the week. For example, given a 10-week M&V
period for 1 intervention, the balanced randomized schedule would
equally sample 5 Mondays with the baseline and 5 Mondays with the
intervention. The limitation of the randomized M&V is that it is only
applicable to a subset of retrofit projects such as control
interventions. However, for all applicable use cases, it allows analysts
to detect energy savings sequentially as the study progresses meaning
once the desired savings target is achieved, analysts can terminate the
M&V. The key advantage of randomization, which is one of the study
objectives, is that if control strategies are sampled with equal
probability, the influence of non-routine events is likely to be evenly
distributed between the baseline and intervention measurements. This
means that the savings estimate, calculated as the difference between
the two, effectively 'cancels out' the impact of these disturbances,
leading to a more accurate and unbiased assessment of the intervention's
impact.

@@

The Building Genome Dataset 2 (BGD2) is an extensive open-access dataset
designed to advance research and development in building energy
efficiency and control strategies acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 500 buildings' metadata and realistic operational
information from across North America and Europe, making it one of the
most comprehensive collections of building-related data available for
scientific use. The dataset includes various commercial building types
such as offices, education facilities, public, and retail buildings, and
provides detailed information on their physical characteristics (e.g.
enregy ratings, heating types and floor area) and hourly measurements of
chilled and hot water, electricity, gas usage as well as site outdoor
weather conditions.

For the purpose of this study, we only queried whole-building hourly
electricity usage. All building names and precise locations were erased
but the climate zone and city name were provided, which are shown in
Figure \@ref(fig:tidysite) and Figure \@ref(fig:messysite). In addition,
the measurements were pre-processed by the authors with timestamp
already converted to local time. We further described the data filtering
process in Section \@ref(method).

@@

### Measurement and verification energy-efficient measures

Most research related to M&V for whole-building approach focuses on the
accuracy of baseline modeling, exploring model performance from simple
regression models to more complex machine learning techniques. One study
reviewed various models suitable for M&V applications as well as
selected input features [@alrobaie_review_2022] and another study
provided a definitive methodology to apply machine learning models for
M&V use cases [@gallagher_development_2018]. In addition, a few studies
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015,
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies made
significant contributions by emphasizing the uncertainty associated with
the model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014, @heo_calibration_2012,
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that baseline model might deteriorate (i.e.
becomes 'stale') over an extended period of pre- and post-analysis.

### Impact of non-routine events on building energy usage

A common non-routine event in energy-saving M&V projects is a change in
occupancy or a significant shift in occupant behavior. These changes can
greatly affect measured energy consumption in buildings and are
typically unrelated to the intervention strategy. For instance, during
the COVID-19 pandemic in 2020, most commercial buildings were
unoccupied, leading to a noticeable drop in energy bills despite no
energy-efficiency measures being implemented [@cai_impact_2024,
@chihib_impact_2021, @gaspar_assessing_2022, @kang_changes_2021].
Furthermore, subsequent research has shown that hybrid working modes,
allowing employees to work remotely, have persisted after the outbreak
of the pandemic [@aksoy_working_2022], adding further complexity to
energy consumption patterns due to evolved occupant behaviors with
higher flexibility [@gui_impact_2021, @mantesi_office_2022,
@xie_does_2021]. One study realized the limitation of current M&V
methods, which only consider adjusting for outdoor weather, is
insufficient and emphasized the importance of requiring matched
comparison groups to control for exogenous factors beyond weather
differences when comparing between baseline and intervention
[@demand_side_analytics_population_2022]. Another type of non-routin
event is filter clogging in air handling units due to particle
accumulation. This can cause supply fans to gradually consume more
energy to maintain required duct static pressure [@feng_newly_2019,
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be underestimated as
increased energy use is incorrectly attributed to the intervention
rather than the mechanical issue.

### Randomized experimental design in other scientific fields

To address the research gap identified in existing M&V studies,
particularly the challenges posed by non-routine events, we applied a
randomized experimental design method---an approach commonly used in
other scientific disciplines. In clinical trials, for instance,
randomization is employed to determine the effect of a medical treatment
(such as a new drug) by randomly assigning participants to either a
control group (e.g., receiving a placebo) or a treatment group
[@angus_adaptive_2019, @burger_importance_2021, @lim_randomization_2019,
@wiley_crossover_2016]. The primary purpose of randomization is to block
confounding variables such as age, gender, and socioeconomic status when
evaluating treatment effects. In the context of buildings, a "treatment"
refers to a control retrofit, while the "placebo" represents the
existing baseline control. Unlike clinical trials, where participants
are randomly assigned, our approach randomizes treatment assignment
longitudinally for each individual building---an approach known as
n-of-1 trials [@gabler2011]. Although the BDG2 dataset allows for
randomization at a population level (e.g., across climate zones),
building owners are generally more interested in understanding the
effects of control retrofits specific to their own buildings, rather
than at a generalized population level.

@@

As mentioned, the goal of an M&V project is to determine the
effect---typically energy savings---of an energy-efficient intervention.
In this study, we focus primarily on switchable interventions, which
often involve control retrofits. An example of such an intervention is a
control retrofit developed by a software-as-a-service company that
adjusts the chilled water plant's supply water temperature based on
outdoor weather conditions [@roa_field_2023,
@lee_simulationoptimization_2012, @qiu_chilled_2022]. Therefore, we
defined the M&V scenario as follows:

*"A company aims to sell its supply temperature reset control software
package to a customer, such as a building owner, with a guarantee that
it will reduce the building's electricity usage. If the building owner
decides to purchase the service, the company agrees to charge a service
fee based on a percentage of the measured energy savings."*

As required by the M&V scenario, we assessed the performance of both the
conventional and the novel randomized M&V methods by estimating the
intervention energy savings for all valid buildings in the dataset. By
conducting such analysis, we aim to:

1.  Compare the energy saving estimation accuracy between the
    conventional and the randomized method. This study extends the
    comparison to a large sample of buildings, covering a variety of
    types and climate zones. The comparison metrics include both
    estimation accuracy and M&V finishing timeline.

2.  Verify the enhanced robustness of the randomized method. By using
    realistic measurements from real-world buildings, which include
    various sources of noise, we aim to reflect the challenges faced by
    building analysts in real projects. As will be demonstrated in the
    following sections, the randomized approach is less impacted by
    non-routine events (i.e., measurement 'noise'), resulting in more
    reliable energy savings estimates.

3.  Demonstrate the implementation of the proposed randomized M&V method
    using a public available dataset. We ensured the reproducibility of
    the method by making the analysis code open source including
    randomized schedule generation, sequential statistical analysis,
    energy modeling and normalized saving calculation. Using the
    available open resources, building analysts should be able to
    seamlessly integrate and apply them in their own M&V projects.

@@

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section.

@@

1.  Missing values \< 1000: given the hourly resolution of all
    measurements, this is equivalent to 1.5 months of missing days.

@@

4.  Target buildings should have known site location: buildings with
    anonymous location are excluded due to unavailable typical
    meteorological weather.

5.  Warehouse and parking types are excluded: target buildings have less
    demand flexibility to implement a chilled water setpoint reset
    control.

Therefor the resulting subset contains all buildings with 'tidy'
measurements. Figure \@ref(fig:tidysite) shows in total, the subset
contains `r nrow(all_names_tidy)` buildings in `r nrow(all_types_tidy)`
types from `r nrow(all_sites_tidy)` different climate zones.

@@

In reality, the 'tidy' subset is less representative of the measurements
typically collected from existing building stock, as whole-building
electricity measurements collected over two years often show more
variability the the filtering critieron. These variations can stem from
issues like sensor calibration errors or inherent changes in the
building, which, as previously discussed, can bias M&V results.
Therefore, to more realistically assess the robustness of the two M&V
method, we included an additional 'messy' subset which first exclude the
'tidy' subset and then re-apply the filtering rule with one amendment:

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease of building electricity usage in the second year should be
    less than 25% of that in th first year.

Figure \@ref(fig:messysite) shows the summary of the 'messy' dataset,
which contains `r nrow(all_names_messy)` buildings in
`r nrow(all_types_messy)` types from `r nrow(all_sites_messy)` different
climate zones.

@@

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that reset the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022, @congradac_recognition_2012]. For both
strategies, we assume that the chiller is activated when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

@@

We mapped the chilled water supply temperature reset to the electrical
energy savings as: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
further consumes around 50% of the HVAC electricity. While this
assumption largely simplifies the diverse energy usage across various
building types, for the scope of this paper, we assume that 25% of the
total building electricity is used by the chilled water plant, $E_{CHW}$
[@us2012commercial]. Typically, the savings from an intervention are not
proportional to the building's hourly electricity usage, which is
generally the challenge for M&V. To address this, we mapped the
resulting electricity savings as a percentage of the plant's normal
operation, calculated as its mean electricity usage over the two-year
period. This percentage is influenced by factors such as outdoor
temperature ($OAT$), intervention supply water temperature
($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
of the day ($\delta_{occ}$, binary indicator whether during peak hours
from 9 AM to 4 PM).

@@

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building, and were used uniformly across the dataset.

|                        |                                                                  |       |
|------------------|------------------------------------|------------------|
| Parameter              | Description                                                      | Value |
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

: Table 1. Parameters for calculating the intervention savings.

@@

To illustrate the impact of non-routine events that are not properly
adjusted for, we developed three scenarios to quantify the energy
consumption changes associated with a hypothetical increase in
occupancy. Specifically, we assume that a rise in occupancy in 2016
(referred to as the "baseline year" when using the conventional method)
led to a 20% increase in whole-building measured electricity usage. The
three scenarios describe when this increase occurred: January to April
(S1), May to August (S2), and September to December (S3). This
hypothetical change was applied only to buildings in the "tidy" subset,
as these buildings were initially more likely to be free from
non-routine events.

@@

As described in Section \@ref(introduction), conventional method for M&V
is time-consuming and unable to separate non-routine event impact from
measured savings. If using conventional method to estimate intervention
energy savings, the result (after 24 months) is calculated by the
difference between measured intervention and projected baseline in the
post-retrofit period. In this study, we leveraged a piece-wise linear
regression considering time-of-week and outdoor temperature (TOWT) as
independent variables for projection for the counter-factual baseline in
the post-retrofit period and normalization on typical meteorological
year [@mathieu_quantifying_2011].

@@

Compared to the conventional method, the randomized M&V approach offers
a more rapid and reliable estimation of energy savings. To apply this
method, analysts first define the target savings and stopping criteria.
Then they design a randomized switchback schedule and perform sequential
statistical tests, such as the sequential probability ratio test (SPRT),
to monitor savings as data is collected. Once the target savings are
detected, the analysts fit a prediction model (e.g. TOWT) to adjust for
differences in outdoor temperature, following the same adjustment
process as the conventional method. We provide example switchback
experimental design and stopping criteria thresholds below:

@@

ASHRAE Guideline 14 offering minimum requirements for whole-building
measurement path states that the baseline period is either a full range
of all independent variables (typically outdoor weather conditions)
under normal facility operation or a 12-month worth of continuous
measurements. The same requirements also applies to intervention
installed in the post-retrofit period. Thus, the conventional M&V method
is likely to run over 24 months or even longer due to missing data. For
example if an analyst is asked to determine the energy savings of a
chilled water plant retrofit but missed most of the cooling season due
to delay in retrofit deployment, he/she needs to wait until the next
cooling season to measure the savings. If the randomized M&V method is
applicable, delays in retrofit deployment pose less risk to the building
owner since baseline are continuously monitored. Sampling at equal
probabilities (e.g., 50%/50% between baseline and intervention) helps to
balance the distribution of independent variables like outdoor weather
conditions across all sampled control strategies. As long as one
strategy is not sampled over a large number of consecutive days (e.g., 7
days or more), it is unlikely that the measurements of independent
variables will differ significantly between the control strategies.

Figure \@ref(fig:timeline) shows the average estimated timeline for each
building across all sites and climate conditions if using randomized M&V
for the applied chilled water supply temperature reset intervention. The
timeline figures for each individual building are attached in the
supplementary material. The figure shows for all climate zones, covering
a sufficient range of outdoor weather condition is the most stringent
requirement. However, most buildings can achieve 80% of the required
range within 6 months, and all buildings can determine the savings,
including associated uncertainties, within 9 months---significantly
shorter than the baseline measurement period required by the
conventional method.

@@

In this study, we define the overall target savings (i.e., the ground
truth) that an M&V analyst aims to detect as the mean reduction in
electricity consumption over a two-year period resulting from the
implementation of the chilled water supply temperature reset control.
This target can be expressed in terms of normalized savings, fractional
savings, or simply the measured difference between baseline and
intervention measurements. Due to different timeline required by the two
methods, we compare the conventional M&V savings at the end of the
two-year period with the randomized M&V savings after all criteria are
satisfied. Since those are the time in reality, an M&V analyst report
estimated savings.

Figure \@ref(fig:norm) shows the overall comparison of the savings
estimated for all buildings in both subsets normalized on the typical
meteorological year weather conditions of each site. In subplot a), the
narrower range and clustering of the true savings indicates the
dependence of the intervention effect on the outdoor weather condition,
which is intended after weather normalization. Sites with mild climate
all year round such as locations in California shows higher savings
potential above 10%, while locations with more extreme climate such as
Washington DC shows only 6% savings annually. In subplot b), the results
indicate that the conventional M&V method tends to estimate savings with
greater uncertainty, and its distribution median deviates from the true
savings. In comparison, the randomized method, which requires much less
time, shows more accurate and precise estimation results.

@@

p1 <- df_FS_tidy %>% 
  bind_rows(df_FS_messy) %>% 
  filter(method == "true") %>% 
  mutate(method = "true savings") %>% 
  ggplot(aes(x = method, y = savings, fill = method)) +
  geom_errorbar(aes(ymin = min(savings), ymax = max(savings)), width = 0.2) +
  geom_boxplot(outlier.size = 2) +
  geom_text(aes(x = 1.35, 
                y = median(savings) + 1.5, 
                label = paste0("median: ", round(median(savings), digits = 0), "%")), 
            check_overlap = T) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     limits = c(0, 30), 
                     labels = number_format(suffix = "%")) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

p2 <- df_FS_tidy %>% 
  bind_rows(df_FS_messy) %>% 
  pivot_wider(names_from = method, values_from = savings) %>% 
  left_join(df_seq_FS_tidy %>% 
              bind_rows(df_seq_FS_messy) %>% 
              filter(seq == "eob"), by = c("name", "site")) %>% 
  pivot_longer(c(conv, rand, FS), names_to = "parameter", values_to = "value") %>% 
  filter(parameter != "rand") %>% 
  mutate(parameter = as.factor(parameter), 
         parameter = recode_factor(parameter, 
                                   "conv" = "Conventional M&V", 
                                   "FS" = "Randomized M&V")) %>% 
  ggplot(aes(x = parameter, y = value, fill = parameter)) +
  stat_boxplot(geom ='errorbar', width = 0.5) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = breaks_pretty(n = 3), 
                     limits = c(0, 30)) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        axis.text.y = element_blank(),
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

  ggarrange(p1, p2,
            labels = c("a)", "b)"),
            ncol = 2, nrow = 1,
            widths = c(0.5, 1),
            legend = "none") +
    plot_annotation(title = "Normalized annual fractional savings estimation comparison", 
                    subtitle = "between convention and randomized M&V (with early stop)")
```

To assess accuracy at the individual building level, we calculated the
mean absolute difference between the true savings and the estimated
savings and normalized the results as a fraction of measured baseline
for generalized comparison. Figure \@ref(fig:tidyfrac) shows the
comparison between the two methods using the tidy subset. In this
comparison, we focused on the median and the 95th percentile of the
plotted distribution to minimize the influence of outliers. The results
indicate that the conventional method outperforms the randomized method
in this subset. This is mostly because the tidy subset includes
buildings with consistent usage patterns throughout the two-year period,
meaning the data used for regression model fitting closely aligns with
the conditions during the model prediction phase. In our case, this
justifies the TOWT model selection when there is no additional
adjustment needed for such pre- and post-analysis.
